---
title: "Results"
author: "Colin Crooks"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document:
    toc: true
    toc_float: true
    fig_caption: true
---




```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE)

###########################################################
#https://github.com/ateucher/useful_code/blob/master/R/numbers2words.r
#https://gist.github.com/psychemedia/numbers2words.R
numbers2words <- function(x){
  ## Function by John Fox found here: 
  ## http://tolstoy.newcastle.edu.au/R/help/05/04/2715.html
  ## Tweaks by AJH to add commas and "and"
  helper <- function(x){
    
    digits <- rev(strsplit(as.character(x), "")[[1]])
    nDigits <- length(digits)
    if (nDigits == 1) as.vector(ones[digits])
    else if (nDigits == 2)
      if (x <= 19) as.vector(teens[digits[1]])
    else trim(paste(tens[digits[2]],
                    Recall(as.numeric(digits[1]))))
    else if (nDigits == 3) trim(paste(ones[digits[3]], "hundred and", 
                                      Recall(makeNumber(digits[2:1]))))
    else {
      nSuffix <- ((nDigits + 2) %/% 3) - 1
      if (nSuffix > length(suffixes)) stop(paste(x, "is too large!"))
      trim(paste(Recall(makeNumber(digits[
        nDigits:(3*nSuffix + 1)])),
        suffixes[nSuffix],"," ,
        Recall(makeNumber(digits[(3*nSuffix):1]))))
    }
  }
  trim <- function(text){
    #Tidy leading/trailing whitespace, space before comma
    text=gsub("^\ ", "", gsub("\ *$", "", gsub("\ ,",",",text)))
    #Clear any trailing " and"
    text=gsub(" and$","",text)
    #Clear any trailing comma
    gsub("\ *,$","",text)
  }  
  makeNumber <- function(...) as.numeric(paste(..., collapse=""))     
  #Disable scientific notation
  opts <- options(scipen=100) 
  on.exit(options(opts)) 
  ones <- c("", "one", "two", "three", "four", "five", "six", "seven",
            "eight", "nine") 
  names(ones) <- 0:9 
  teens <- c("ten", "eleven", "twelve", "thirteen", "fourteen", "fifteen",
             "sixteen", " seventeen", "eighteen", "nineteen")
  names(teens) <- 0:9 
  tens <- c("twenty", "thirty", "forty", "fifty", "sixty", "seventy", "eighty",
            "ninety") 
  names(tens) <- 2:9 
  x <- round(x)
  suffixes <- c("thousand", "million", "billion", "trillion")     
  if (length(x) > 1) return(trim(sapply(x, helper)))
  helper(x)
  }
  
  rnd <- function(x) floor(x/5)*5
  
  median.iqr <- function(x, dig=0) paste0(round(median(x, na.rm = T))," (",paste(round(quantile(x,c(0.25,0.75),na.rm = T),digits = dig),collapse = ","),")")
  
  n.perc <- function(x, dig=0) paste0(rnd(sum(x == T, na.rm = T))," (",round(mean(x,na.rm = T),digits = dig)*100,"%)")
  
  date.90.day <- function(x,ref.dat) is.finite(x) & as.numeric(x) - as.numeric(ref.dat) <= 90  & as.numeric(x) - as.numeric(ref.dat) >=0
  
  safelog <- function(x) { x[x < 1e-200] <- 1e-200; log(x) }
  
  cuminc.km <- function(x,niter)  { 
    n.type.events <- sort(unique(dt.tv[(postop.covid.cohort),event]))[-1]
    data.table::setkey(dt.tv, patient_id, tstart,tstop)
    est <- boot.est <- Reduce(merge,lapply(n.type.events, function(i) data.table::setDT(summary(survival::survfit(survival::Surv(start,end,event==i) ~ get(x), 
                                                                                 data = dt.tv[(postop.covid.cohort)], 
                                                                                 id = patient_id),
                                                               times = sort(unique(dt.tv[(postop.covid.cohort) & event == i,end]))
    )[c('strata','time','n.risk', 'n.event')])[,(paste0('haz',i)) := n.event/n.risk][,c(1,2,5)])
    )[order(strata,time),
      .(time,cumsum(exp(cumsum(log(1 - Reduce('+',.SD))))*haz1)),
      keyby = strata, 
      .SDcols = paste0('haz',n.type.events)][order(strata,time),
                                   round(tail(V2,1),digits = 3)*100, 
                                   keyby = strata] 
  
  for(iter in 1:niter) {
    samp.patient_id <- dt.tv[patient_id %in% sort(sample(unique(dt.tv[(postop.covid.cohort),patient_id]), replace = T))]
    data.table::setkey(dt.tv, patient_id, tstart,tstop)
    boot.est <- cbind(boot.est,
                      Reduce(merge,
                             lapply(n.type.events,
                                    function(i) {
       event.times = sort(unique(dt.tv[(postop.covid.cohort) & event == i,end]))
       if(samp.patient_id[(event==i),.N] > 0) {
          data.table::setDT(summary(survival::survfit(survival::Surv(start,end,event==i) ~ get(x), 
      data = samp.patient_id,
      id = patient_id),
      times = event.times,
      extend = T)[c('strata','time','n.risk', 'n.event')])[
        , (paste0('haz',i)) := n.event/n.risk][,c(1,2,5)]
       }
       else {
        data.table::data.table('strata' = rep(levels(est[[1]]), 
                                              each = length(event.times)),
                               'time' = rep(event.times, 
                                            times = length(levels(est[[1]]))))[,(paste0('haz',i)) := 0] 
       }
     }
    )
)[order(strata,time),.(time,
                       cumsum(exp(cumsum(safelog(1 - Reduce('+',.SD))))*haz1)),
  keyby = strata, 
  .SDcols = paste0('haz',n.type.events)][order(strata,time),
                                         round(tail(V2,1),digits = 3)*100,
                                         keyby = strata][,2])
}
  return(cbind(est[,1],
               rnd(dt.tv[(postop.covid.cohort) & tstop == final.date,.N, keyby = get(x)][,2]),
               rnd(data.table::setDT(summary(survival::survfit(survival::Surv(start,end,event==1) ~ get(x), 
                   data = dt.tv[(postop.covid.cohort),],
                   id = patient_id), times = 90)[c('n.event')])),
               est[,2],
               t(apply(boot.est[,2:(iter + 1)],1,quantile,c(0.025,0.975), na.rm = T))))
  }

##########################################################


#########################
# Basic counts and descriptions from raw data
#############################
index_date <- data.table::as.IDate("2020-02-01")
dt <- data.table::fread(here::here("output", "input.csv"))

procedures <- c('Abdominal','Cardiac','Obstetrics','Orthopaedic','Thoracic', 'Vascular')
procs <- paste0(rep(procedures,each = 5),"_",1:5)

dt[,dateofbirth := (data.table::as.IDate(paste0(dob,'-15')))]
dt[dereg_date != "",gp.end := data.table::as.IDate(paste0(dereg_date,'-15'))]
dt[, imd := as.numeric(imd)]
dt[, imd5 := cut(imd, breaks = seq(0,33000,33000/5),  include.lowest = T, ordered_result = F)]

####################################################################
# Multiple operations per row. Reshape to long format
####################################################################

repeated.vars <- names(dt)[grepl(pattern = paste(procedures,collapse = '|'),x = names(dt))]
non.op.vars <- names(dt)[!grepl(pattern = paste(procedures,collapse = '|'),x = names(dt))]

aggregate_operations <- data.table::melt(dt, id.var = 'patient_id',
                                         measure = patterns(as.vector(outer(paste0(procedures,'_[0-9]'),
                                                                            unique(gsub(pattern = paste0(as.vector(outer(procedures,paste0("_",1:5),paste0)),collapse = "|"),replacement = "",x = repeated.vars)),
                                                                            paste0))),
                                         value.name = as.vector(outer(procedures,unique(gsub(pattern = paste0(as.vector(outer(procedures,paste0("_",1:5),paste0)),collapse = "|"),replacement = "",x = repeated.vars)),paste0)))
dt <- dt[,non.op.vars, with = F][aggregate_operations, on = "patient_id"]
rm(aggregate_operations)
dt[, keep := F]
dt[!is.na(dereg_date), dereg_date := data.table::as.IDate(paste0(dereg_date,"-30"), format = "%Y-%m-%d")]
for (x in paste0(procedures,"_date_admitted")) { dt[is.na(dereg_date) | x <= dereg_date, keep := T] }
dt <- dt[keep == T,]

n.ops <- rnd(sum(unlist(lapply(paste0(procedures,"_date_admitted"), function(x) dt[is.finite(get(x)),.N]))))
n.pats <- rnd(sum(length(unique(dt[,patient_id]))))
start.date <- data.table::as.IDate(min(vapply(paste0(procedures,"_date_admitted"), 
                                              FUN.VALUE = 1.0,
                                              FUN =function(x) unlist(dt[is.finite(get(x))][order(get(x)),get(x), with = T][1]))))
last.date <- data.table::as.IDate(max(vapply(paste0(procedures,"_date_admitted"), 
                                             FUN.VALUE = 1.0, 
                                             FUN =function(x) unlist(dt[is.finite(get(x))][order(-get(x)),get(x), with = T][1]))))

dt[,(paste("admit.wave.",procedures, sep ="")) := lapply(.SD, function(x) cut(as.numeric(x),
                breaks =
                  c(as.numeric(data.table::as.IDate("2020-01-01", format = "%Y-%m-%d")), 
                    as.numeric(data.table::as.IDate("2020-09-01", format = "%Y-%m-%d")),                                       
                    as.numeric(data.table::as.IDate("2021-05-01", format = "%Y-%m-%d")),
                    as.numeric(data.table::as.IDate("2021-12-31", format = "%Y-%m-%d")),
                    as.numeric(data.table::as.IDate("2022-05-01", format = "%Y-%m-%d"))),
                    labels = c("Wave_1","Wave_2","Wave_3","Wave_4"),
                    ordered = T)), 
   .SDcols = c(paste0(procedures,"_date_admitted"))]
for(x in procedures) {
  dt[, (paste0(x,"post.VTE")) := ((!is.na(.SD[,3]) &  
                                     .SD[,3] <= .SD[,1] + 90 & .SD[,3] >= .SD[,1]) | 
                                    ((!is.na(.SD[,4]) &  
                                        .SD[,4] <= .SD[,1] + 90 & .SD[,4] >= .SD[,1]))) & 
       (!is.na(.SD[,5]) &  
          .SD[,5] <= .SD[,1] + 90 & .SD[,5] >= .SD[,1]), 
     .SDcols = paste0(procedures,c("_date_admitted","_date_discharged",
                                   "_VTE_GP_date","_VTE_HES_date_admitted",
                                   "_anticoagulation_prescriptions_date"))]  # events flagged at end of episode
}

```

## Study population
One third of GP practices in England use SystmOne, with records for approximately 44% of the English population. In this study `r numbers2words(n.pats)` adult patients were identified undergoing at least one procedure whilst registered to a general practitioner using the SystmOne database at the time of the first procedure between `r format(start.date, format="%B %d %Y")` and `r format(last.date, format="%B %d %Y")`. `r stringr::str_to_sentence(numbers2words(n.ops))` procedures were identified within thse patients' registration period after restricting to the first 5 procedures within each category (abdominal, cardiac, obstetric, orthopaedic,  thoracic, & Vascular). The numbers and demographics of these patients and procedures are shown in Table 1.

```{r demo, results='asis', echo = F, warning=F}
load(file = here::here("output","cohort_long.RData"))
data.table::setkey(dt.tv,patient_id,tstart,tstop)
dt.tv[,Emergency := max(Emergency, na.rm = T), by = .(patient_id, end.fu)]

data.table::setkey(dt.tv,patient_id,tstart,tstop)
dt.tv[,bmi := max(bmi, na.rm = T), by = .(patient_id, end.fu) ]
dt.tv[,imd5 := max(as.numeric(imd5), na.rm = T), by = .(patient_id, end.fu) ]
dt.tv[!is.finite(vaccination.status), vaccination.status := 0]
dt.tv[,vaccination.status := cumsum(vaccination.status), by = .(patient_id, end.fu) ]
dt.tv[!is.finite(Current.Cancer), Current.Cancer := 0]
dt.tv[,Current.Cancer := cumsum(Current.Cancer), by = .(patient_id, end.fu) ]
dt.tv[!is.finite(Emergency), Emergency := 0]
dt.tv[,Emergency := cumsum(Emergency), by = .(patient_id, end.fu) ]
dt.tv[!is.finite(died), died := 0]
dt.tv[,died := cumsum(died), by = .(patient_id, end.fu) ]
dt.tv[!is.finite(postcovid), postcovid := 0]
dt.tv[,postcovid := cumsum(postcovid), by = .(patient_id, end.fu) ]
dt.tv[!is.finite(postVTEany), postVTEany := 0]
dt.tv[,postVTEany := cumsum(postVTEany), by = .(patient_id, end.fu) ]
dt.tv[,region := max(region, na.rm = T), by = .(patient_id, end.fu) ]
dt.tv[,primary_diagnosis := max(primary_diagnosis, na.rm = T), by = .(patient_id, end.fu) ]
dt.tv[start<=90 & end>90, end := 90]
dt.tv <- dt.tv[end <= 90]

demo.tab <- data.table::transpose(cbind(
dt.tv[tstop == end.fu & !is.na(op.type),.('Procedures' = rnd(.N),
                         'Patients' = rnd(length(unique(patient_id))), 
                         'Female' = n.perc(sex=='F',dig=3),
                         'Age (IQR)' = median.iqr(age,dig = 0),
                         "BMI (IQR)" =  median.iqr(bmi,dig = 0),
                         "IMD quintile (IQR)" = median.iqr(as.numeric(imd5),dig = 0),
                         "1st Vaccination" = n.perc(vaccination.status == 1,dig = 3),
                         "2nd Vaccination" = n.perc(vaccination.status == 2,dig = 3),
                         "3rd Vaccination" = n.perc(vaccination.status == 3,dig = 3),
                         "Current Cancer"  = n.perc(Current.Cancer == T,dig = 3),
                         "Emergency" = n.perc(Emergency == T,dig = 0),
                         "90 day mortality (%)" =n.perc(died == T,dig = 4),
                         "90 day COVID-19 (%)" = n.perc(postcovid == 1, dig = 3),
                         "90 day VTE (%)" = n.perc(postVTEany == 1, dig = 3)), keyby = op.type],
<<<<<<< Updated upstream
data.table::transpose(
  cbind(
    data.table::transpose(
      data.table::dcast(
        cbind(
          dt.tv[tstop == end.fu & !is.na(op.type),
                rnd(table(op.type)),
                keyby = region],
          rep(procedures,times = 10)),
        V2 ~ region, 
        value.var = 'V1' ),
      keep.names = "region", 
      make.names = "V2")[,
                         lapply(.SD,
                                function(x) paste0(x, 
                                                   ' (',
                                                   round(100*x/sum(x,na.rm = T), 
                                                         digits = 1),
                                                   '%)')
                                ),
                         .SDcols = 2:7],
    sort(unique(dt.tv$region))
    ),
  make.names = "V2"),
data.table::transpose(
  cbind(1:5,
        data.table::transpose(
          data.table::dcast(
            dt.tv[tstart == admit.date & !is.na(op.type),
                  rnd(.N),
                  keyby = c('op.type','primary_diagnosis')],
            op.type ~ primary_diagnosis,
            value.var = 'V1'), 
          keep.names = "Primary_Diagnosis",
          make.names = "op.type")[,1:7][,
                                        lapply(.SD, 
                                               function(x) paste0(Primary_Diagnosis,
                                                                  ": ",
                                                                  x, 
                                                                  ' (',
                                                                  round(100*x/sum(x,na.rm = T),
                                                                        digits = 1),
                                                                  '%)')[order(x)]), 
                                        .SDcols = 2:7][1:5,]), 
  make.names = 'V1')
),
make.names = 'op.type',
keep.names = 'Characteristics')
=======
data.table::transpose(cbind(data.table::transpose(data.table::dcast(cbind(dt.tv[tstop == end.fu & !is.na(op.type),rnd(table(op.type)), keyby = region],rep(procedures,times = 10)), V2 ~ region, value.var = 'V1' ),keep.names = "region", make.names = "V2")[,lapply(.SD, function(x) paste0(x, ' (',round(100*x/sum(x,na.rm = T), digits = 1),'%)')), .SDcols = 2:7],sort(unique(dt.tv$region))), make.names = "V2"),
data.table::transpose(cbind(1:5,data.table::transpose(data.table::dcast(dt.tv[tstart == admit.date & !is.na(op.type),rnd(.N), keyby = c('op.type','primary_diagnosis')],op.type ~ primary_diagnosis, value.var = 'V1'), keep.names = "Primary_Diagnosis", make.names = "op.type")[,1:7][,lapply(.SD, function(x) paste0(Primary_Diagnosis,": ", x, ' (',round(100*x/sum(x,na.rm = T), digits = 1),'%)')[order(x)]), .SDcols = 2:7][1:5,]), make.names = 'V1')),make.names = 'op.type', keep.names = 'Characteristics')
>>>>>>> Stashed changes
# 
#                           
# 
# demo.tab <- xtable::xtable(
#   data.table::rbindlist(
#     list(lapply(procedures, function(proc) { 
#   t(cbind(dt[is.finite(get(paste0(proc,'_date_admitted'))) ,.("Procedures" = rnd(.N),
#                                  "Patients" = rnd(length(unique(patient_id))),
#                                  "Female" = n.perc(sex=='F',dig = 3),
#                                  "Age (IQR)" = median.iqr(as.numeric(get(paste0(proc,'_date_admitted')) - as.numeric(dateofbirth))/365.25,dig = 0),
#                                  "BMI (IQR)" =  median.iqr(bmi,dig = 0),
#                                  "IMD quintile (IQR)" = median.iqr(as.numeric(imd5),dig = 0),
#                                  "1st Vaccination" = n.perc(is.finite(covid_vaccine_dates_1) & get(paste0(proc,'_date_admitted'))  - covid_vaccine_dates_1 >= 14,dig = 3),
#                                  "2nd Vaccination" = n.perc(is.finite(covid_vaccine_dates_2) & get(paste0(proc,'_date_admitted'))  - covid_vaccine_dates_2 >= 14,dig = 3),
#                                  "3rd Vaccination" = n.perc(is.finite(covid_vaccine_dates_3) & get(paste0(proc,'_date_admitted'))  - covid_vaccine_dates_3 >= 14,dig = 3),
#                                  "Current Cancer"  = n.perc(substr(get(paste0(proc,'_primary_diagnosis')),1,1) =='C',dig = 3),
#                                  "Emergency" = n.perc(substr(get(paste0(proc,'_admission_method')),1,1) == "2",dig = 3),
#                                  "Length of stay (IQR)" =  median.iqr(get(paste0(proc,'_date_discharged')) - get(paste0(proc,'_date_admitted')),dig = 0),
#                                  "90 day mortality (%)" =n.perc(date.90.day(x = date_death_ons, ref.dat = get(paste0(proc,'_date_admitted'))),dig = 4),
#                                  "90 day COVID-19 (%)" = paste0(rnd(sum(get(paste0(proc,'_case_category')) != "", na.rm = T))," (", round(100*mean(get(paste0(proc,'_case_category')) != "", na.rm = T), digits = 1),"%)"),
#                                  n.perc(((date.90.day(x = get(paste0(proc,"_VTE_HES_date_admitted")), ref.dat = get(paste0(proc,'_date_admitted')))) |
#                                  (date.90.day(x = get(paste0(proc,"_VTE_GP_date")), ref.dat = get(paste0(proc,'_date_admitted'))))) &
#                                  date.90.day(x = get(paste0(proc,"_anticoagulation_prescriptions_date")), ref.dat = get(paste0(proc,'_date_admitted'))),dig = 4))],
#                                t(dt[is.finite(get(paste0(proc,'_date_admitted'))) ,rnd(.N), keyby = region][,paste0(V1, ' (',round(100*V1/sum(V1,na.rm = T), digits = 1),'%)')]),
#                                t(dt[is.finite(get(paste0(proc,'_date_admitted'))),rnd(.N), by = .(.grp = eval(parse(text = paste0(proc,'_primary_diagnosis'))))][order(-V1), paste0(do.call(paste,c(.SD, sep = ": ")),' (',round(100*V1/sum(V1,na.rm = T),digits = 1),'%)')][1:5]))
#     )
#       }
#   )
#   )
#   ),include.rownames = T)
# 
# 
# rownames(demo.tab) <- c("Procedures (N)","Patients (N)","Female (%)","Median Age, years (IQR)","Median BMI (IQR)","IMD quintile (IQR)","1st Vaccination (%)","2nd Vaccination (%)","3rd Vaccination (%)","Cancer diagnosis (%)", "Emergency (%)","Median length of stay (IQR)","90 day mortality (%)","90 day COVID-19 (%)","90 day VTE (%)",unique(dt$region), "Top 5 primary diagnoses: 1", "2","3","4","5")
# colnames(demo.tab) <- procedures
                        
knitr::kable(demo.tab, caption = "Demographics of patients by procedure type with counts rounded to nearest 5")

```


## Post operative 90 day COVID-19 risk
Table 2 shows the cumulative risk of testing positive for COVID-19 peri-opertively for elective patients, from 7 days prior to the operation (i.e within pre operative quarantine periods for elective procedures) to 90 days post operatively. 

```{r cuminc, results='asis', echo = F, warning=FALSE, eval=FALSE}
procedures <- unique(dt.tv$op.type)


data.table::setkey(dt.tv,patient_id,tstart,tstop)
dt.tv[,final.date := covid.end]
dt.tv[is.finite(readmit.end) & readmit.end < final.date, final.date := readmit.end]
dt.tv[is.finite(end.fu) & end.fu < final.date, final.date := end.fu]

dt.tv[,event :=0]
dt.tv[COVIDpositive == T, event := 1]
dt.tv[emergency_readmit == T & COVIDpositive !=T, event := 2]
dt.tv[died == T & COVIDpositive !=T, event := 3]

dt.tv[, postop.covid.cohort := start>=0 & tstop <= final.date & end <= 90]


covariates <- c('age.cat','sex','op.type','Charl12','bmi.cat','imd5','region',
                'vaccination.status.factor','Emergency','Current.Cancer','wave')

data.table::setkey(dt.tv,patient_id,tstart,tstop)

crude.covid.cov <- data.table::rbindlist(lapply(1:length(covariates), function(i) cbind(rep(covariates[i],length(levels(dt.tv[(postop.covid.cohort),as.factor(get(covariates[i]))]))),
                                    levels(dt.tv[(postop.covid.cohort),as.factor(get(covariates[i]))]),
                                    cuminc.km(covariates[i], niter = 1000)[,2:6])))

names(crude.covid.cov) <- c("Characteristic","Level","Number at risk",
                            "Number of events","90 day Cumulative Risk",
                            "Bootstrapped Confidence","95% interval" )
knitr::kable(crude.covid.cov, caption = "Demographics of patients by procedure type with counts rounded to nearest 5")

```